
\section{Contributors}\label{contributors}

\paragraph{Authors}
\begin{itemize}
    \item Team Wizkit: Paul Bunn, David Darais, Daniel Genkin, Steve Lu, Kimberlee Model, Tarik Riviere, Muthuramakrishnan Venkitasubramaniam, Xiao Wang
    \item Team Emphasize:
    Steven Eker, Karim Eldefrawy, Stéphane Graham-Lengrand, Vitor Pereira, Hadas Zeilberger 
    \item Team Fromager: Michael Adjedj, Daniel Benarroch Guenun, Eran Tromer, Aurélien Nicolas, Constance Beguier
    \item Team Oracles: Mayank Varia
\end{itemize}

The following individuals from the T\&E Team were instrumental in moderating the IR's development:
David A. Wilson (PI),
R. Nicholas Cunningham,
J. Parker Diamond,
Hanson Duan,
Ariel Hamlin,
Noah Luther,
Richard Shay.\\

\noindent\textbf{Disclaimers}
\\

This material is based upon work supported by DARPA under Contracts No.~HR001120C0087, HR001120C0086, HR001120C0085 and Agreement No.~HR00112020021.  Any opinions, findings and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of DARPA.\\

\textbf{Distribution Statement A:} Approved for public release. Distribution Unlimited.\\

The authors of the SIEVE IR have made every effort that the SIEVE IR Specification, if adopted by a recognized standards body, would be compatible with \href{https://opensource.org/osr}{the OSI Open Standard Requirement}.

\begin{description}
  \item[No Intentional Secrets] We have made every effort to ensure that the IR contains no necessary secrets. Further, several of the SIEVE Performers have or will open-source their IR implementations.
  \item[Availability] The SIEVE Performers have agreed to release the text of the IR Specification under the \href{https://creativecommons.org/licenses/by/4.0/}{Creative Commons Attribution 4.0 International (CC-BY 4.0)} terms.
  \item[Patents] Although we are not aware of any current patents which may be necessary to implement the IR, the SIEVE Performers have \href{https://github.mit.edu/sieve-all/collaboration/issues/30}{agreed not to enforce} any intellectual property rights which may be necessary to implement the IR.
    [ When we do release the Spec, I can also take a print-out of this issue. ]
  \item[Dependencies] To our knowledge, all technologies dependent by the IR are already themselves open-source.
\end{description}
