\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage[export]{adjustbox}
\usepackage[title,titletoc]{appendix}

\usepackage{amssymb,amsmath}
\usepackage{hyperref}
\usepackage{xcolor}

\usepackage{tabularx}
\usepackage{cleveref}
\usepackage{ltablex}
\usepackage{alltt}
\usepackage{listings}
\usepackage{enumitem}

\definecolor{InternalLinkColor}{HTML}{282888}
\definecolor{ExternalLinkColor}{HTML}{3333BB}

\hypersetup{
  colorlinks=true,
  linkcolor=InternalLinkColor,
  urlcolor=ExternalLinkColor
}

\definecolor{light-gray}{gray}{0.95}

\lstdefinelanguage{ir}
{
  backgroundcolor=\color{light-gray},
  morekeywords={version, circuit, public_input, private_input, configuration, field, @field, ext_field, ring, @begin, @end, @function, @out, @in, @public, @private, @call, @anon_call, @type, @plugin},
  morekeywords=[2]{@add, @addc, @mul, @mulc, @assert_zero, @delete, @convert, @new},
  keywordstyle={[2]\it},
  morecomment=[l]{//}
}

\lstdefinelanguage{flatbuffer}
{
  backgroundcolor=\color{light-gray},
  morekeywords={table,struct,union},
  morekeywords=[2]{uint32,uint64,ubyte,string},
  keywordstyle={[2]\it}
}

\lstdefinelanguage{json}{
    backgroundcolor=\color{light-gray},
    basicstyle=\ttfamily\small,
}

\lstset{escapechar=\#}

\definecolor{SyntaxGreen}{HTML}{116611}
\definecolor{SemanticPurple}{HTML}{550077}

\renewcommand{\familydefault}{\sfdefault}

\title{Introduction to the SIEVE Intermediate Representation}
\author{See Section \ref{contributors} for list of \nameref{contributors}}
\date{Last Updated: 2023-07-17}

\newcommand{\maxfieldcount}{256}

\begin{document}
\maketitle

\textbf{Distribution Statement ``A'':} Approved for Public Release, Distribution Unlimited.\\

This material is based upon work supported by DARPA under Contracts No.~HR001120C0087, HR001120C0086, HR001120C0085 and Agreement No.~HR00112020021.  Any opinions, findings and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of DARPA.

\section*{Identification and History}\label{history}

\begin{tabularx}{\textwidth}{|p{1in}|p{1in}|p{4in}|}
  \hline
  \textbf{Version} & \textbf{Date} & \textbf{Notable Changes}                                                                                                                                                                   \\
  \hline
  0.1.0            & 2020-09-19    & Initial version as proposed jointly by WizKit and FROMAGER teams on DARPA SIEVE.                                                                                                           \\
  \hline
  0.1.1
                   & 2020-12-16    & A number of bugfixes on the prior version.
  \begin{itemize}
    \item Fix typographical errors and omissions.
    \item Add checks for matching header, \texttt{num\_wires}, and etc., across multiple resources.
    \item Eliminate unused type-checking and arity-checking attributes from the attribute grammar, because IR0 has only one meaningful type.
    \item Add \texttt{num\_wires} to the instance and witness in the text and binary grammars.
  \end{itemize}                                                                                       \\
  \hline
  0.2.0            & 2021-03-19    & Remove the \texttt{num\_wires} and associated input size data and add a \texttt{@delete} directive.                                                                                        \\
  \hline
  1.0.0            & 2021-07-20    & Add ``uniformity'' features: function gates, public-index for loops, and private-condition switch-case statements.                                                                         \\
  \hline
  1.0.1            & 2022-04-29    & Fix the following issues:
  \begin{itemize}
    \item Fix switch-statement semantics which required a nested switch-statement where the outer case is inactive and the inner switch-statement has no active branch to cause an unexpected witness-statement invalidity.
    \item Constrain for-loop bounds and wire-ranges such that the first must be less than or equal to the last iteration or wire.
    \item Add a warning that the switch-statement conversion algorithm may produce a poorly-formed relation with certain forms of switch-statement and for-loop nesting. Add a suggestion to unroll for-loops in those conditions.
    \item Assorted typo fixes and performance improvements.
  \end{itemize} \\
  \hline
  2.0.0            & 2022-11-08    & The 2.0.0 revision is a significant departure from prior revisions, regressing some of the haphazardly developed high level features and preparing for the development of a distinct high-level (translation), and low-level (circuit) IR layers.
  This release was used during the SIEVE Phase II testing event, and specifies the Circuit-IR -- characterized as follows.
  \begin{itemize}
    \item Flat circuit with numbered wires.
    \item Function's encapsulating circuit fragments.
    \item Explicit allocation directives enable memory efficiency.
    \item A new plugin system allows access to backend-defined functionality.
  \end{itemize} \\
  \hline
  2.1.0            & 2023-07-17    & Adds ring and extension field types to the Circuit-IR and multi-wire convenience modes for copy gates and input gates. \\
  \hline
\end{tabularx}

\newpage
\tableofcontents
\newpage

%%%%
\input{01-intro}
%%%%
\input{02-headers}
%%%%
\input{03-subdocuments}
%%%%
\input{07-contributors}

\end{document}
